---
title: "newrmarkdownfinaltemplatefordata"
output: html_document
author: Kylie Peasley
date: "2025-10-30"
---



``` {r eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 #(data set will be imported here)
```
# ABSTRACT

# BACKGROUND



# Choose some colors
cols <- c(rainbow(30)[c(10:29,1:5)])  # rainbow colors

# horizontal barplot
par(mar=c(5,12,2,2))  # wider left margin for names
bp <- barplot(duration$mean, horiz=TRUE, names.arg=duration$Bird,
              las=1, col=cols, xlab="Days of detectable viremia", xlim=c(0,7))

# add error bars
arrows(duration$lo, bp, duration$hi, bp,
       angle=90, code=3, length=0.05, col="black", xpd=TRUE)
```
```{r eval=FALSE}
#the data will be imported once I figure out how to do that
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

#basic barplot
barplot(height=data$value, names=data$name)
```

# STUDY QUESTION and HYPOTHESIS

## Questions  
 Does the presence of music decrease the sensation of pain?

## Hypothesis 
  Music therapy decreases the sensation of pain compared to no music or pink noise.
## Prediction


# METHODS


##1st analysis (barplots)

##2nd analysis (generalized linear regression model)


```{r test name}
# second-analysis-or-plot, music presence against binary +/_
#This test will determine if there is a significant relationship between whether a patient is listening to music or not and their pain levels reported
# second-analysis-or-plot, music presence against binary +/_
# Placeholder data for demonstration
data <- data.frame(
  pain_present = sample(0:1, 20, replace = TRUE),   # binary outcome
  music_present = sample(c(0, 1), 20, replace = TRUE),  # binary predictor
  pain_score = runif(20, 0, 10)  # continuous variable (e.g., pain rating)
)

# Test if music presence predicts pain presence
glm1 <- glm(pain_present ~ music_present,
            data = data,
            family = binomial)
summary(glm1)

# GLM with music presence against pain score
glm2 <- glm(pain_score ~ music_present,
            data = data)
summary(glm2)

```

# DISCUSSION
 

## Interpretation of 1st analysis (barplots)



## Interpretation of 2nd analysis (generalized linear model)


# CONCLUSION



# REFERENCES     
1. 

2. ChatGPT. OpenAI, version Jan 2025. Used as a reference for functions such as plot() and to correct syntax errors. Accessed `r Sys.Date()`.

